Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/RelatedWork.tex

\section{Related work}
\label{relatedWork}
Formal verification of cache coherence protocols, among other hardware systems,
is mainly done using model checking. Explicit state model checking tools like
Murphi \cite{murphi} or TLC \cite{tlc} have been used in practise only to
handle small and simplified systems, \eg{} a single level cache hierarchy with
three CPUs handling two addresses and one bit of data. Symbolic model checking
techniques have fared slightly better, McMillan \etal{} have verified a 2-level
MSI protocol based on Gigamax distributed multiprocessor using SMV.
Optimizations on these techniques (like partial order reduction \cite{part},
symmetry reduction \cite{sym1, sym2}, compositional reasoning \cite{somethink}, \etc
scale the model checking approach further, but they still haven't been able to
verify realistic hierarchical coherence protocols.

Theorem prover based approaches have also been used to verify cache coherence
protocols. FLASH coherence protocol \cite{flash} has been verified using an
approach called Aggregation of Distributed transactions \cite{park}.  Several
state-transitions in the implementation accessing localized states intuitively
correspond to a single atomic transition accessing all the states globally. In
this approach, the user has to provide a proof of a mapping between the
implemantion transitions to the aggregated atomic transition; once this is
done, a model checker is used to verify the aggregated transition resulting in
verification of the implementation. We believe that the difficult part in
proving a protocol is precisely the proof of such a mapping from an
implementation-level transition to a specification-level transitions accessing
states globally, and their work does not present any approach to go about it.

>>>>

\section{Related work}
\label{relatedWork}

Formal verification of hardware-oriented cache coherence protocols, is mainly
done using model checking. In practice, explicit state model checking tools like
Murphi \cite{murphi} or TLC \cite{tlc} are only able to handle small and
simplified systems, \eg{} a single-level cache hierarchy with three CPUs, two
addresses, and one bit of data. Symbolic model checking techniques have fared
slightly better; McMillan \etal{} have verified a 2-level MSI protocol based on
Gigamax distributed multiprocessor using SMV.  Optimizations on these techniques
(\eg{} partial order reduction \cite{part}, symmetry reduction \cite{sym1,
  sym2}, compositional reasoning \cite{somethink}, \etc{} scale the approach
further, but they are still unable to handle realistic hierarchical protocols.

Theorem provers have also been used in verifying cache coherence protocols. The
FLASH coherence protocol \cite{flash} was verified using Aggregation of
Distributed transactions \cite{park}, where several state transitions accessing
localized state intuitively correspond to an aggregate atomic transition
accessing all global state. A proof of a mapping between the local transitions
and the agreggate one must be provided. Once this is done, a model checker can
be used to verify the aggregate model is correct. However, We believe that the
verification challenge here is precisely the proof of such a mapping which was
not discussed.
